---
title: "Analysis of Tesla Stock Data "
author: "Group 7"
date: "12/12/2023"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Introduction

This report presents a detailed analysis of Tesla's stock data. The focus is to understand the trends and patterns in the stock prices, particularly looking at volume traded over time, and to forecast future movements.

## Data Loading and Preprocessing

The Tesla stock data was loaded and preprocessed, focusing on formatting the dates and converting numeric values for in-depth analysis. The following code was used for this purpose:

```{r, echo=FALSE, fig.height=3}
# Loading the libraries
library(ggplot2)
library(forecast)
library(tseries)
library(tseries)
library(TSA)
library(readxl)
library(xts)

# Loading the data
df <- read.csv("/Users/hongzifei/Downloads/TeslaStockData.csv", header = TRUE)

df$Date <- as.Date(df$Date)   # Converting 'Date' to Date format


df <- df[complete.cases(df$Date), ]
df_xts <- xts(df[, -1], order.by = df$Date)  # Converting to xts time series
# Assuming numeric columns are from the 2nd column onwards
num_cols <- 2:6
df[, num_cols] <- lapply(df[, num_cols], function(x) as.numeric(gsub("[\\$,]", "", x)))


# Converting to xts time series with numeric columns
df_xts <- xts(df[, num_cols], order.by = df$Date)

# Obtaining monthly average
df_monthly <- as.data.frame(apply.monthly(df_xts, colMeans))
df_monthly$Date <- index(df_monthly)  # Add the 'Date' column back

head(df_monthly)  # Displaying the first few rows
summary(df_monthly)  # Summary statistics
```

##Exploratory Data Analysis
In this phase, we explored the data, focusing on the Volume traded over time.
```{r, echo=FALSE}
# Boxplots for each variable - possible indicator for extreme values/ peaks in volume
par(mfrow=c(2,2))

boxplot(df_monthly$Volume, main="Boxplot - Volume")

#time series plots

ggplot(df_monthly, aes(x = Date, y = Volume)) + geom_line() + labs(title = "Time Series Plot - Monthly Average Volume")
```

##Time Series Decomposition
We decomposed the time series data to analyze the trend, seasonality, and noise components.
```{r, echo=FALSE, fig.height=3}
# Decomposition of the time series plots (additive)
decomp_additive <- decompose(ts(df_monthly$Volume, frequency = 12), type = "additive")
plot(decomp_additive)

# Decomposition of the time series plot (multiplicative)

decomp_multiplicative <- decompose(ts(df_monthly$Volume, frequency = 12), type = "multiplicative")
plot(decomp_multiplicative)
```

##Autocorrelation Analysis
Autocorrelation and partial autocorrelation plots were used to understand the correlations in the data.
```{r, echo=FALSE, fig.height=3}
# Autocorrelation plots (ACF)

acf(df_monthly$Volume, lag.max=40)

# Partial autocorrelation plots (PACF)

pacf(df_monthly$Volume, lag.max=40)
```

##Stationarity Testing and Transformation
Stationarity of the data was tested to ensure the validity of time series models.
```{r, echo=FALSE, fig.height=3}
# ADF test and Log Transformations
## Data Splitting

# Assuming your data is already sorted by date
cutoff_index <- round(0.7 * nrow(df_monthly))
train_data <- df_monthly[1:cutoff_index, ]
test_data <- df_monthly[(cutoff_index + 1):nrow(df_monthly), ]

# ADF test on original and transformed data
adf.test(train_data$Volume)
plot(log(df_monthly$Volume), type = "l")
plot(diff(log(df_monthly$Volume)), type = "l")
adf.test(log(df_monthly$Volume))
adf.test(diff(log(df_monthly$Volume)))

```

##SARIMA Model Fitting and Forecasting
We fitted a SARIMA model to the data and made future forecasts.
```{r, echo=FALSE, fig.height=3}

# Fitting the SARIMA model
arima_model_volume <- auto.arima(train_data$Volume)
sarima_model_volume <- Arima(train_data$Volume, order = c(1, 0, 0), seasonal = list(order = c(1, 1, 1), period = 12))
print(summary(sarima_model_volume))

# Forecasting
sarima_forecast <- forecast(sarima_model_volume, h = length(test_data$Volume))

# Creating the sequence for the forecasted values
forecast_sequence <- seq(from = length(train_data$Volume) + 1, 
                         to = length(train_data$Volume) + length(test_data$Volume))

# Plotting the original data, test data, and forecast
plot(df_monthly$Volume, type = "l", col = "blue", lwd = 2, ylab = "Volume", xlab = "Time", main = "SARIMA Forecast vs. Original Data")
lines(forecast_sequence, test_data$Volume, col = "green", lwd = 2)
lines(forecast_sequence, sarima_forecast$mean, col = "red", lwd = 2)
legend("topright", legend = c("Original Data", "Test Data", "Forecast"), col = c("blue", "green", "red"), lty = 1:2, cex = 0.8)

```

# Ljung-Box Test
```{r, echo=FALSE}
sarima_residuals <- residuals(sarima_model_volume)
ljung_box_test <- Box.test(sarima_residuals, lag = 40, type = "Ljung-Box")
print(ljung_box_test)
```


